The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-06 manuscript/manuscript.Rnw
fix how gene.expression is called cgb.min.expresison

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    alignment.species <- species.name
    if (alignment.species=="pan paniscus") {
        alignment.species <- "pan troglodytes"
    } else if (alignment.species=="ateles fusciceps") {
        alignment.species <- "homo sapiens"
    } else if (grepl("spalax",alignment.species)) {
        alignment.species <- "mus musculus"
    }
    temp <- 
        combined.fpkm[species==species.name &
                      !(!is.na(human_name) & 
                        human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                      ! (gene_biotype %in% c("Mt_tRNA","Mt_rRNA","rRNA")) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?$)",human_name) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?$)",human_alignment_symbol) &
                      !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[l]?(?:-.+)?$|rRNA$)",
                             gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_alignment_symbol,
                                                 human_name,mean_fpkm)]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    temp <- temp[!duplicated(name_or_id)][1:n]
    temp[,used_alignment:=FALSE]
    temp[is.na(human_name) & !is.na(human_alignment_symbol),
                    used_alignment:=TRUE]
    temp[is.na(human_name),human_name:=human_alignment_symbol]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,used_alignment,`Approved Symbol`,`Approved Name`,mean_fpkm)]
    temp[!is.na(name_or_id),
         name_or_id:=paste0("\\href{http://www.ensembl.org/",
                             capfirst(gsub(" ","_",alignment.species)),
                             "/Gene/Summary?g=",name_or_id,"}{",
                             xtable.sanitize.text(name_or_id),"}")]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Symbol`),"}")]
    temp[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
    temp[,used_alignment:=NULL]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM"))
    if (species.name=="homo sapiens" || species.name=="ateles fusciceps") {
        temp[,`Human Symbol`:=NULL]
    }
    return(temp)
}
gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE) {
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm]
    fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\mathrm{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0("\\mathrm{",gene,"}",
                           fpkm)
        } else {
            fpkm <- paste0("\\mathit{",gene,"}$=$",
                           fpkm)
        }
    }
    if (units || complete) {
        fpkm <- paste0("$",fpkm,"$")
    }
    return(fpkm)
}

@@

  <<cgb.min.expression,results="hide",echo=FALSE,warnings=FALSE>>=
cgb.min.expression <- 
    min(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x)}))
cgb.sum.expression <- 
    sum(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x)}))
@

>>>>

<<top.genes.function,echo=FALSE,results="hide">>=
top.genes.table <- function(species.name,n=50) {
    alignment.species <- species.name
    if (alignment.species=="pan paniscus") {
        alignment.species <- "pan troglodytes"
    } else if (alignment.species=="ateles fusciceps") {
        alignment.species <- "homo sapiens"
    } else if (grepl("spalax",alignment.species)) {
        alignment.species <- "mus musculus"
    }
    temp <- 
        combined.fpkm[species==species.name &
                      !(!is.na(human_name) & 
                        human_name %in% housekeeping.genes.superset[,gene_short_name]) &
                      ! (gene_biotype %in% c("Mt_tRNA","Mt_rRNA","rRNA")) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?$)",human_name) &
                      !grepl("((RNA\\d|MT)-|^RP[SL]P?\\d+[ABCD]?$)",human_alignment_symbol) &
                      !grepl("(^[Mm][Tt]-|^R[pP][lsLS]\\d+[abcdABCD]?[l]?(?:-.+)?$|rRNA$)",
                             gene_short_name)
                      ][order(-mean_fpkm)][,list(name_or_id,
                                                 human_alignment_symbol,
                                                 human_name,mean_fpkm)]
    ## remove any genes which have a null name_or_id; they are probably from unassembled contigs
    temp <- temp[!(is.na(name_or_id)|name_or_id==""),]
    temp <- temp[!duplicated(name_or_id)][1:n]
    temp[,used_alignment:=FALSE]
    temp[is.na(human_name) & !is.na(human_alignment_symbol),
                    used_alignment:=TRUE]
    temp[is.na(human_name),human_name:=human_alignment_symbol]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,used_alignment,`Approved Symbol`,`Approved Name`,mean_fpkm)]
    temp[!is.na(name_or_id),
         name_or_id:=paste0("\\href{http://www.ensembl.org/",
                             capfirst(gsub(" ","_",alignment.species)),
                             "/Gene/Summary?g=",name_or_id,"}{",
                             xtable.sanitize.text(name_or_id),"}")]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Symbol`),"}")]
    temp[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
    temp[,used_alignment:=NULL]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM"))
    if (species.name=="homo sapiens" || species.name=="ateles fusciceps") {
        temp[,`Human Symbol`:=NULL]
    }
    return(temp)
}
gene.expression <- function(species.name,gene,complete=FALSE,units=TRUE) {
    fpkm <- combined.fpkm[species==species.name &
                              (gene_short_name==gene | 
                               gene_id==gene),
                          mean_fpkm]
    if (units || complete)
        fpkm <- formatC(fpkm,digits=0,format="f")
    if (units) {
        fpkm <- paste0(fpkm,
                       "\\ \\mathrm{FPKM}")
    }
    if (complete) {
        if (any(grepl("^ENS.+G\\d+$",gene))) {
            fpkm <- paste0("\\mathrm{",gene,"}",
                           fpkm)
        } else {
            fpkm <- paste0("\\mathit{",gene,"}$=$",
                           fpkm)
        }
    }
    if (units || complete) {
        fpkm <- paste0("$",fpkm,"$")
    }
    return(fpkm)
}

@@

<<cgb.min.expression,results="hide",echo=FALSE,warnings=FALSE>>=
cgb.min.expression <- 
    min(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x,units=FALSE)}))
cgb.sum.expression <- 
    sum(sapply(paste0("CGB",c("","1","2","5","7","8")),
               function(x){gene.expression(species.name="homo sapiens",gene=x,units=FALSE)}))
@
